---
title: "Introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Introduction}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
options(rmarkdown.html_vignette.check_title = FALSE)
```


This package `moewishart` provides miximum likelihood estimation (MLE) and Bayesian estimation for the **Wishart mixture model** and **Wishart mixture-of-experts** (**MoE-Wishart**) model. 
It contains 4 different algorithms for 2 model:

- mixture model of Wishart distributions:
    - EM algorithm for finding the MLE;
    - Bayesian approach using Gibbs-within-MH sampling algorithm.
- Mixture-of-Expert model, in which the probability gating can depend on covariate(s):
   - EM-MoE algorithm for finding the MLE;
   - Bayesian-MoE approach using Gibbs-within-MH sampling algorithm.  


## Installation

Install the latest development version from [GitHub](https://github.com/zhizuio/moewishart)

```{r eval=FALSE}
#library("devtools")
devtools::install_github("zhizuio/moewishart", ref="main", auth_token = "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")
```

## Example


<br> 

Data simulation from a MoE-Wishart model:

- Sample size $n = 200$
- Dimension of the Wishart distribution $p = 2$
- Number of latent components $K = 3$
- $q$ covariates $\mathbf X = \{x_{ij}\} \in \mathbb R^{n\times q}$, $x_{ij}\sim\text{N}(0,1)$
- Fixed covariate effects $\boldsymbol\beta \in \mathbb R^{q\times K}$, with $\boldsymbol\beta_{K}=0$
- Probabilities of subpopulations $\boldsymbol\pi = \{\pi_{ik}\} \in \mathbb R^{n\times q}$, $\pi_{ik} = \exp\{\mathbf X_i\boldsymbol\beta_k\} / \sum_{l=1}^K\exp\{\mathbf X_i\boldsymbol\beta_k\}$
- Degrees of freedom $\boldsymbol\nu = (8, 12, 3)$
- Scale matrices of the Wishart distribution $\Sigma_1$, $\Sigma_3$, $\Sigma_3$
- Data $S_i \sim \pi_1\text{Wishart}(\nu_1, \Sigma_1) + \pi_2\text{Wishart}(\nu_2, \Sigma_2) + \pi_3\text{Wishart}(\nu_3, \Sigma_3)$


### 1. Working model: Bayesian MoE-Wishart model

```{r, results='hide'}
library(moewishart)

set.seed(123)
n <- 200 # number of subjects
p <- 2 # dimension of covariance matrix
set.seed(123) # fix coefficients of underlying MoE model
Xq <- 3; K = 3
betas <- matrix(runif(Xq * K, -2, 2), nrow = Xq, ncol = K)
betas[, K] <- 0

# simulate data
dat <- simData(n, p,
  Xq = 3, K = 3, betas = betas,
  pis = c(0.35, 0.40, 0.25),
  nus = c(8, 12, 3)
)

# fit Bayesian MoE-Wishart model
set.seed(123)
fit <- moewishartX(
  dat$S, X = cbind(1, dat$X), K = 3, 
  mh_sigma = c(0.2, 0.1, 0.2), # RW-MH variances (length K)
  mh_beta = c(0.3, 0.3), # RW-MH variances (length K-1)
  niter = 3000, burnin = 1000
)
```

<br> 

```{r}
# posterior summaries
apply(fit$nu, 2, mean) # estimated nu
dat$nu # true nu

# Wishart scale matrices
MoE_Sigma <- Reduce("+", fit$Sigma) / length(fit$Sigma)
MoE_Sigma

# gating coefficients
burnin <- 1000
beta_mcmc <- fit$Beta_samples[-c(1:burnin), , ]
apply(beta_mcmc, c(2, 3), mean)
```


### 2. Working model: Bayesian Wishart mixture model

```{r, results='hide'}
# fit Bayesian Wishart mixture model
set.seed(123)
fit2 <- moewishart(
  dat$S, K = 3, 
  mh_sigma = c(0.2, 0.1, 0.2), # RW-MH variances
  niter = 3000, burnin = 1000
)
```

<br> 

```{r}
# posterior summaries
colMeans(fit2$pi[-c(1:burnin), ]) # estimated pis

colMeans(fit2$nu[-c(1:burnin), ]) # estimated nu
```


### 3. Working model: MoE-Wishart model via EM algorithm

```{r}
# fit MoE-Wishart model via EM alg.
set.seed(123)
fit3 <- moewishartX(
  dat$S, X = cbind(1, dat$X), K = 3, 
  method = "em",
  niter = 3000
)
```

<br> 

```{r}
# EM estimates
fit3$nu

fit3$Sigma

fit3$Beta
```


### 4. Working model: Wishart mixture model via EM algorithm

```{r}
# fit Wishart mixture model via EM alg.
set.seed(123)
fit4 <- moewishart(
  dat$S, K = 3, 
  method = "em",
  niter = 3000
)
```

<br> 

```{r}
# EM estimates
fit4$nu

fit4$Sigma
```


## References

The Tien Mai, Zhi Zhao (2026). xxxx

