---
title: "Simulation study: mixture of Wishart"
author: "The Tien Mai & Zhi Zhao"
date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document
---

```{css, echo=FALSE}
pre code {
  white-space: pre-wrap;
}
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "", eval = TRUE)
options(width = 140)
```

<br> 

Data simulation:

- Sample size $n = 200$
- Dimension of the Wishart distribution $p = 10$
- Number of latent components $K = 5$
- Probabilities of subpopulations $\boldsymbol\pi = (0.35, 0.10, 0.25, 0.2, 0.1)$
- Degrees of freedom $\boldsymbol\nu = (p+5, p+2, p+10, p+1, p+3)$
- Scale matrices of the Wishart distribution $\Sigma_k$, $k=1,...,K$
- Data $S_i \sim \sum_{k=1}^K\pi_k\text{Wishart}(\nu_k, \Sigma_k)$

Modeling:

- Bayesian mixture of five Wishart distributions



```{r}
rm(list = ls())

library(moewishart)

set.seed(123)
n <- 200 # number of subjects
p <- 10

dat <- simData(n, p,
  K = 5,
  pis = c(0.35, 0.10, 0.25, 0.2, 0.1),
  nus = c(p + 5, p + 2, p + 10, p + 1, p + 3),
  Sigma = NULL # default Sigmas pre-defined in function 'simData()'
)
S_list <- dat$S
Sigma_list <- dat$Sigma_list

my_K_intial <- 5

# run a moewishart model
start_time <- Sys.time()
set.seed(123)
fit <- moewishart(S_list,
  K = my_K_intial, mh_sigma = 1,
  niter = 10000, burnin = 1000, thin = 1, verbose = TRUE
)
end_time <- Sys.time()
elapsed_time <- end_time - start_time
elapsed_time
```

<br> 

```{r}
# posterior summaries
colMeans(fit$pi)
colMeans(dat$pi) # true pi

apply(fit$nu, 2, mean)
dat$nu # true nu
```


<br> 

## EM algorithm

<br> 

```{r}
# run the model with EM algorithm
set.seed(1234)
fit3 <- moewishart(S_list,
  method = "em",
  K = my_K_intial, mh_sigma = 1,
  niter = 10000, burnin = 1000, thin = 1, verbose = TRUE
)

# EM estimates
fit3$pi
colMeans(dat$pi) # true pi

fit3$nu
dat$nu # true nu
```




