---
title: "Simulation study: mixture of experts model with covariates"
author: "The Tien Mai & Zhi Zhao"
date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document
---

```{css, echo=FALSE}
pre code {
  white-space: pre-wrap;
}
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "", eval = TRUE)
options(width = 140)
```

<br> 

Data simulation:

- Sample size $n = 200$
- Dimension of the Wishart distribution $p = 2$
- Number of latent components $K = 3$
- $q$ covariates $\mathbf X = \{x_{ij}\} \in \mathbb R^{n\times q}$, $x_{ij}\sim\text{N}(0,1)$
- Fixed covariate effects $\boldsymbol\beta \in \mathbb R^{q\times K}$, with $\boldsymbol\beta_{K}=0$
- Probabilities of subpopulations $\boldsymbol\pi = \{\pi_{ik}\} \in \mathbb R^{n\times q}$, $\pi_{ik} = \exp\{\mathbf X_i\boldsymbol\beta_k\} / \sum_{l=1}^K\exp\{\mathbf X_i\boldsymbol\beta_k\}$
- Degrees of freedom $\boldsymbol\nu = (8, 12, 3)$
- Scale matrices of the Wishart distribution $\Sigma_1$, $\Sigma_3$, $\Sigma_3$
- Data $S_i \sim \pi_1\text{Wishart}(\nu_1, \Sigma_1) + \pi_2\text{Wishart}(\nu_2, \Sigma_2) + \pi_3\text{Wishart}(\nu_3, \Sigma_3)$

Modeling:

- Bayesian mixture of experts of two Wishart distributions



```{r, results='hide'}
rm(list = ls())

library(mewishart)

set.seed(123)
n <- 200 # number of subjects
p <- 2

dat <- simData(n, p,
  Xq = 3, K = 3,
  nus = c(8, 12, 3),
  Sigma = NULL # default Sigmas pre-defined in function 'simData()'
)
S_list <- dat$S
Sigma_list <- dat$Sigma_list

# run a mewishart model with covariates
start_time <- Sys.time() # this has to be before 'set.seed()', otherwise non-reproducible
set.seed(123)
fit <- mewishartX(S_list, K = 3, X = dat$X)
end_time <- Sys.time()
elapsed_time <- end_time - start_time
```

<br> 

```{r}
str(fit)
# posterior summaries

colMeans(fit$pi) # estimated pis
colMeans(dat$pi) # true pis

apply(fit$nu, 2, mean) # estimated nu
dat$nu # true nu

# Sum all arrays in the list and divide by the number of saved iterations
MoE_mean <- Reduce("+", fit$Sigma) / length(fit$Sigma)
# View the mean covariance matrix for Cluster 1
print(MoE_mean[, , 1])
print(MoE_mean[, , 2])
print(MoE_mean[, , 3])
as.data.frame(Sigma_list)
(a2 <- apply(MoE_mean, MARGIN = 3, function(aa) min(sapply(Sigma_list, function(x) sum((aa - x)^2)))))
```


