---
title: "Simulation study: mixture of experts model"
author: "The Tien Mai & Zhi Zhao"
date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document
---

```{css, echo=FALSE}
pre code {
  white-space: pre-wrap;
}
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "", eval = TRUE)
options(width = 140)
```

<br> 

Data simulation:

- Sample size $n = 200$
- Dimension of the Wishart distribution $p = 2$
- Number of latent components $K = 3$
- Probabilities of subpopulations $\boldsymbol\pi = (0.35, 0.40, 0.25)$
- Degrees of freedom $\boldsymbol\nu = (8, 12, 3)$
- Scale matrices of the Wishart distribution $\Sigma_1$, $\Sigma_3$, $\Sigma_3$
- Data $S_i \sim \pi_1\text{Wishart}(\nu_1, \Sigma_1) + \pi_2\text{Wishart}(\nu_2, \Sigma_2) + \pi_3\text{Wishart}(\nu_3, \Sigma_3)$

Modeling:

- Bayesian mixture of experts of two Wishart distributions



```{r}
rm(list = ls())

library(moewishart)

set.seed(5)
n <- 200 # number of subjects
p <- 2

dat <- simData(n, p,
  pis = c(0.35, 0.40, 0.25),
  nus = c(8, 12, 3),
  Sigma = NULL # default Sigmas pre-defined in function 'simData()'
)
S_list <- dat$S
Sigma_list <- dat$Sigma_list

my_K_intial <- 3

# run a moewishart model
start_time <- Sys.time() # this has to be before 'set.seed()', otherwise non-reproducible
set.seed(123)
fit <- moewishart(S_list,
  K = my_K_intial, init_pi = c(0.45, 0.40, 0.15), mh_sigma = 1,
  niter = 10000, burnin = 1000, thin = 1, verbose = TRUE
)
end_time <- Sys.time()
elapsed_time <- end_time - start_time
elapsed_time
```

<br> 

```{r}
# posterior summaries
colMeans(fit$pi) # estimated pi
colMeans(dat$pi) # true pi

apply(fit$nu, 2, mean) # estimated nu
dat$nu # true nu

# Sum all arrays in the list and divide by the number of saved iterations
sigma_posterior_mean <- Reduce("+", fit$Sigma) / length(fit$Sigma)
# View the mean covariance matrix for Cluster 1
print(sigma_posterior_mean[, , 1])
print(sigma_posterior_mean[, , 2])
print(sigma_posterior_mean[, , 3])
as.data.frame(Sigma_list)

(a1 <- apply(sigma_posterior_mean, MARGIN = 3, function(aa) min(sapply(Sigma_list, function(x) sum((aa - x)^2)))))
```

## Check identifiability issue

```{r}
# re-run the model with wrong initial weights to check identifiability issue
set.seed(1234)
fit2 <- moewishart(S_list,
  K = my_K_intial, init_pi = c(0.15, 0.40, 0.45), mh_sigma = 1,
  niter = 10000, burnin = 1000, thin = 1, verbose = TRUE
)

colMeans(fit2$pi) # estimated pi
colMeans(dat$pi) # true pi

apply(fit2$nu, 2, mean) # estimated nu
dat$nu # true nu

# Sum all arrays in the list and divide by the number of saved iterations
sigma_posterior_mean2 <- Reduce("+", fit2$Sigma) / length(fit2$Sigma)
# View the mean covariance matrix for Cluster 1
print(sigma_posterior_mean2[, , 1])
print(sigma_posterior_mean2[, , 2])
print(sigma_posterior_mean2[, , 3])
as.data.frame(Sigma_list)
```


<br> 

## EM algorithm

<br> 

```{r}
# run the model with EM algorithm
set.seed(123)
fit3 <- moewishart(S_list,
  method = "em",
  K = my_K_intial, #init_pi = c(0.33, 0.33, 0.34),
  niter = 10000, verbose = TRUE
)

# EM estimates
fit3$pi
colMeans(dat$pi) # true pi

fit3$nu
dat$nu # true nu
```


