---
title: "Simulation study: mixture of experts model to check identifiability"
author: "The Tien Mai & Zhi Zhao"
date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document
---

```{css, echo=FALSE}
pre code {
  white-space: pre-wrap;
}
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "", eval = TRUE)
options(width = 140)
```

<br> 

Data simulation:

- Sample size $n = 200$
- Dimension of the Wishart distribution $p = 2$
- Number of latent components $K = 2$
- Probabilities of subpopulations $\boldsymbol\pi = (0.8, 0.2)$
- Degrees of freedom $\boldsymbol\nu = (3, 12)$
- Scale matrices of the Wishart distribution $\Sigma_1$, $\Sigma_2$
- Data $S_i \sim \pi_1\text{Wishart}(\nu_1, \Sigma_1) + \pi_2\text{Wishart}(\nu_2, \Sigma_2)$

Modeling:

- Bayesian mixture of experts of two Wishart distributions



```{r}
rm(list = ls())

library(mewishart)

set.seed(123)
n <- 200 # number of subjects
p <- 2

dat <- simData(n, p,
  pis = c(0.8, 0.2),
  nus = c(3, 12),
  Sigma = NULL # default Sigmas pre-defined in function 'simData()'
)
S_list <- dat$S
Sigma_list <- dat$Sigma_list

my_K_intial <- 2

# run a mewishart model
set.seed(123)
fit <- mewishart(S_list,
  K = my_K_intial, init_pi = c(0.9, 0.1),
  niter = 10000, burnin = 1000, thin = 1, verbose = TRUE
)

# posterior summaries
colMeans(fit$pi)
colMeans(dat$pi) # true pi

apply(fit$nu, 2, mean)
dat$nu # true nu
```

<br> 

## Check identifiability issue

```{r}
# run the model again with swapped initial weights
set.seed(123)
fit2 <- mewishart(S_list,
  K = my_K_intial, init_pi = c(0.1, 0.9),
  niter = 10000, burnin = 1000, thin = 1, verbose = TRUE
)

# posterior summaries
colMeans(fit2$pi)
colMeans(dat$pi) # true pi

apply(fit2$nu, 2, mean)
dat$nu # true nu
```

**Conclusion**: 

The two model fittings above with initial weights `init_pi=c(0.9, 0.1)` and `init_pi=c(0.1, 0.9)` result in similar results, which indicate possibly identifiability. 

Based on the proof of Proposition 2 in Yakowitz and Spragins (1968), we can probably give similar proofs to show identifiability of the MoE of (inverse) Wishart distributions.


