---
title: "Simulation study: mixture of experts model"
author: "The Tien Mai & Zhi Zhao"
date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document
---

```{css, echo=FALSE}
pre code {
  white-space: pre-wrap;
}
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "", eval = TRUE)
options(width = 140)
```

<br> 

Data simulation:

- Sample size $n = 200$
- Dimension of the Wishart distribution $p = 2$
- Number of latent components $K = 3$
- Probabilities of subpopulations $\boldsymbol\pi = (0.35, 0.40, 0.25)$
- Degrees of freedom $\boldsymbol\nu = (8, 12, 3)$
- Scale matrices of the Wishart distribution $\Sigma_1$, $\Sigma_3$, $\Sigma_3$
- Data $S_i \sim \pi_1\text{Wishart}(\nu_1, \Sigma_1) + \pi_2\text{Wishart}(\nu_2, \Sigma_2) + \pi_3\text{Wishart}(\nu_3, \Sigma_3)$

Modeling:

- Bayesian mixture of experts of two Wishart distributions



```{r, results='hide'}
rm(list = ls())

library(mewishart)

set.seed(123)
n <- 200 # number of subjects
p <- 2

dat <- simData(n, p,
  pis = c(0.35, 0.40, 0.25),
  nus = c(8, 12, 3),
  Sigma = NULL
)
S_list <- dat$S
Sigma_list <- dat$Sigma_list

# run a mewishart model
start_time <- Sys.time() # this has to be before 'set.seed()', otherwise non-reproducible
set.seed(123)
fit <- mewishart(S_list,
  K = 3, niter = 2000, burnin = 500, thin = 1,
  alpha = rep(.5, p), nu0 = 5, Psi0 = diag(p),
  init_nu = c(8, 12, 10), sample_nu = TRUE,
  nu_prior_a = 3, nu_prior_b = 0.1, mh_sigma = 0.15,
  verbose = TRUE
)
end_time <- Sys.time()
elapsed_time <- end_time - start_time
```

<br> 

```{r}
# posterior summaries
colMeans(fit$pi)
apply(fit$nu, 2, mean)
# Sum all arrays in the list and divide by the number of saved iterations
sigma_posterior_mean <- Reduce("+", fit$Sigma) / length(fit$Sigma)
# View the mean covariance matrix for Cluster 1
print(sigma_posterior_mean[, , 1])
print(sigma_posterior_mean[, , 2])
print(sigma_posterior_mean[, , 3])
Sigma_list
```

```{r}
# Mixture of experts model with covariates
res_MoE <- mewishartX(S_list, X = matrix(rnorm(n * 3), ncol = 3), K = 3)
str(res_MoE)
# posterior summaries
colMeans(res_MoE$pi)
apply(res_MoE$nu, 2, mean)
# Sum all arrays in the list and divide by the number of saved iterations
sigma_posterior_mean <- Reduce("+", res_MoE$Sigma) / length(res_MoE$Sigma)
# View the mean covariance matrix for Cluster 1
print(sigma_posterior_mean[, , 1])
print(sigma_posterior_mean[, , 2])
print(sigma_posterior_mean[, , 3])
as.data.frame(Sigma_list)
```


